---
title: "HW4"
output: html_document
date: "2025-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
```

## Problem 1

```{r problem 1}
nflip(n = 2021, prob = 0.024)

sim_flags <- do(100000) * nflip(n = 2021, prob = 0.024)
head(sim_flags)

ggplot(sim_flags) + 
  geom_histogram(aes(x = nflip), binwidth = 1, fill = "lightblue", color = "black") +
  labs(title = "Distribution of Flagged Trades Under the Null Hypothesis",
       x = "Number of Flagged Trades",
       y = "Frequency")

sum(sim_flags$nflip >= 70)

sum(sim_flags$nflip >= 70) / 100000
```

#### **Null Hypothesis (H₀):**

The proportion of flagged trades for Iron Bank employees is the same as the baseline rate of 2.4%.

#### **Test Statistic:**

Our test statistic is the number of flagged trades out of 2021. Higher numbers show a stronger evidence against the null hypothesis, saying that Iron Bank employees may be flagged at a higher rate.

#### P value:

```         
0.00205
```

#### **Conclusion:**

The p-value is small, so we reject the null hypothesis, assuming that the flagged trade rate for Iron Bank is unusually high and not due to random variation.

## Problem 2

```{r problem 2}
nflip(n = 50, prob = 0.03)

violations <- do(100000) * nflip(n = 50, prob = 0.03)
head(violations)

ggplot(violations) + geom_histogram(aes(x = nflip), binwidth = 1, fill = "green", color = "black") +
  labs(title = "Distribution of Health Code Violations Under Null Hypothesis",
       x = "Number of Violations",
       y = "Frequency")

sum(violations$nflip >= 8)

sum(violations$nflip >= 8) / 100000

```

#### **Null Hypothesis (H₀):**

The proportion of health code violations for Gourmet Bites is the same as the baseline of 3%.

#### **Test Statistic:**

Here, our test statistic is the number of health code violations out of 50 inspections. Higher numbers show stronger evidence against the null hypothesis, conveying that Gourmet Bites might have a higher violation rate.

#### P value:

```         
0.00016
```

#### **Conclusion:**

The p-value is small, so we reject the null hypothesis, suggesting that Gourmet Bites has a higher health violation rate than the city average.

## Problem 3:

```{r problem 3}
expected_proportions <- c(0.30, 0.25, 0.20, 0.15, 0.10)
observed_counts <- c(85, 56, 59, 27, 13)

total_jurors <- sum(observed_counts)
expected_counts <- total_jurors * expected_proportions


chisq.test(x = observed_counts, p = expected_proportions)
```

#### **Null Hypothesis:** 

The judge’s jury selection process reflects the county’s population distribution.

#### **Test Statistic:** 

We use the chi-squared statistic, which measures how different the observed jury composition is from the expected distribution. A larger chi-squared value means a greater deviation from expectation.

#### P Value:

```         
0.01445
```

#### Conclusion:

The observed jury selection is the same as the expected distribution, meaning there is no statistical evidence of bias.

## Problem 4:

### Part A

```{r problem 4}

library(tidyverse)


letter_frequencies = read.csv("letter_frequencies.csv")


letter_frequencies$Probability = letter_frequencies$Probability / sum(letter_frequencies$Probability)


sentences = readLines("brown_sentences.txt")


calculate_chi_squared = function(sentence, freq_table) {
  
 
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  

  clean_sentence = gsub("[^A-Za-z]", "", sentence)
  clean_sentence = toupper(clean_sentence)
  
  
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  
 
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  
  # Compute chi-squared statistic
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)
}


chi_sq_values = rep(0, length(sentences))
for (i in seq_along(sentences)) {
  chi_sq_values[i] = calculate_chi_squared(sentences[i], letter_frequencies)
}

hist(chi_sq_values, breaks = 50, col = "blue",
     main = "Chi-Squared Statistics for Brown Corpus",
     xlab = "Chi-Squared Value", ylab = "Frequency")

```

### Part B

```{r part b}
test_sentences = c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)


test_chi_sq = rep(0, length(test_sentences))
for (i in 1:length(test_sentences)) {
  test_chi_sq[i] = calculate_chi_squared(test_sentences[i], letter_frequencies)
}


test_results = data.frame(Sentence = 1:10, ChiSquared = test_chi_sq)


print(test_results)

LLM_sentence = which.max(test_results$ChiSquared)
cat("The sentence most likely by an LLM is sentence number:", LLM_sentence, "\n")


hist(test_results$ChiSquared, breaks = 5, col = "red",
     main = "Chi-Squared Values for Test Sentences",
     xlab = "Chi-Squared Value", ylab = "Frequency")
```
